---
title: "HW 4"
author: "Eric Zhao"
date: "2023-02-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Uniform random variates


### set-up myunif

```{r}
myunif<-function(n, a, b){
  uni_variates<-rep(NA, n)
  for(i in 1:n){
    uni_variates[i]<-runif(1)*(b-a)+a
  }
  return(uni_variates)
}
```

### simulated results

```{r}
set.seed(1111)
uni_dat1<-myunif(1000, 0, 3)
hist(uni_dat1,freq = FALSE, breaks = "fd")
curve(dunif(x, 0, 3), add = TRUE)
mean(uni_dat1)
sd(uni_dat1)
```
The mean and standard deviation are consistent with the theoretical values $\mu=\frac{3-0}{2}=1.5$ and $\sigma=\frac{3}{\sqrt{12}}=0.866025$.


## Triangular Random variates 


### set-up mytriang

```{r}
mytriang<-function(n, a, b, c){
  tri_variates<-rep(NA, n)
  for(i in 1:n){
    call<-runif(1)
    if(call<(c-a)/(b-a)){
    tri_variates[i]<-sqrt(call*(b-a)*(c-a))+a
    }
    else{
      tri_variates[i]<-b-sqrt((1-call)*(b-a)*(b-c))
    }
  }
  return(tri_variates)
  
}
```

### simulated results

```{r}
set.seed(1031)

tri_dat1<-mytriang(10000, 10, 25, 15)
hist(tri_dat1,freq = FALSE, breaks = "fd")
segments(x0=c(10, 15), y0=c(0, 2/15), x1=c(15, 25), y1=c(2/15, 0))
mean(tri_dat1)
sd(tri_dat1)

```

The mean and standard deviation are consistent with the theoretical values: $\mu=(10+25+15)/3=16.6667$ and $\sigma=\sqrt{\frac{1}{18}(10^2+25^2+15^2-10\times25-10\times15-15\times25)}=3.11805$.





## Normal random variates


### Sum of Uniform Variates Method

#### set-up sumuni-approch of generating normal variate

```{r}
sumuni<- function(n=12){
    z<-(sum(myunif(12, 0, 1))-n/2)/sqrt(n/12)
    return(z)
}
```

### generate normal variates and plot the simulation

```{r}
set.seed(020131)
norm_variates<-rep(NA, 10000)
for (i in 1:10000) {
  norm_variates[i]<-sumuni(n=12)
}

hist(norm_variates,freq = FALSE, breaks = "fd")
curve(dnorm(x), add = TRUE)
```

The normal distribution fit looks reasonable after superimpose the theoretical standard normal curve to it. The range of possible values under this seed is $[-3.718, 3.720]$.


### Box-Muller Method


### BMM function

```{r}
BMM_norm <- function() {
  call1<-myunif(1, 0, 1)
  call2<-myunif(1, 0, 1)
  z1<-sqrt(-2*log(call1))*cos(2*pi*call2)
  z2<-sqrt(-2*log(call1))*sin(2*pi*call2)
  pair<-c(z1, z2)
  return(pair)
}
```


### BMM normal variates simulation

```{r}
set.seed(021031)
norm_variates<-rep(NA, 10000)
for (i in 1:10000) {
  norm_variates[i]<-BMM_norm()[1]
  norm_variates[i+1]<-BMM_norm()[2]
}

hist(norm_variates,freq = FALSE, breaks = "fd")
curve(dnorm(x), add = TRUE)
```

The standard normal variates that generated by this method also looks reasonable but this approach seems to be more efficient than the last sum of uniform variates approach. The possible range for z-score is around $[-4, 4]$($[-3.9054, 4.1996]$ in this case).




